How is experience transformed into memory?

### 2018-08-29 paper/boneyard.tex

>>>>

These features of real world memory provide a stark contrast with classic list-learning memory tasks that ask participants to reproduce from memory a set of studied items (e.g. words)-- whereby each word's meaning is distinct from surrounding words, and successful memory for a given word entails reproducing it exactly during recall.  So how might one go about assessing the quality of real world memory?  Which analysis methods and findings from the list-learning literature apply to real world learning?  And how might one go about formally characterizing whether (or how well) the ``gist'' of a memory has been accurately remembered?

A defining characteristic of

One approach to studying memory for naturalistic experiences and real life events in the laboratory has been to quantify the participant's ability to recall specific aspects of the experience, such as the people, places, and things encountered~\citep[e.g.\ for review see][]{KoriGold94}.  This essentially

Human memory studies typically assess memory for the elements within an experience, such as the people, places and things encountered. However, real life experiences can also be described in terms of their temporal structure (i.e., how the contents relate/change over time).  For instance, naturalistic experiences are autocorrelated in space and time on short timescales. In other words, the contents of an experience often change gradually and are therefore similar from moment to moment. As an example, many of us spend our mornings/evenings at home and much of our days at work (e.g., in an office). While the specific happenings within these contexts differ, they are often highly correlated (and thus predictable) within an experience. Furthermore, our experiences can also be characterized by longer timescale correlations. For example, consider returning to your office after taking a lunch break. The spatial contexts before and after lunch are highly similar, but separated in time. Thus, while evaluating the ability to recall the isolated elements of an experience has been a fruitful way to learn about human memory, another important dimension to consider is the temporal structure of the remembered experience, and how the temporal structure of the memory relates to the original experience. These recurrent and gradually changing temporal dynamics are not typically present in traditional memory studies, but are likely critical to understanding how our memory systems encode, store and retrieve our life experiences.

### 2018-08-29 paper/main.tex

\subsection*{Modeling the naturalistic stimulus and verbal recall}
We fit a topic model \citep{BleiEtal03} to hand-annotated text descriptions of scenes from the video. The text descriptions contained details of the scene such as the characters present, location, and a short summary of the scene (see Fig.~\ref{fig:schematic} for example text). As depicted in Fig.~\ref{fig:model}A, the topic vectors are sparse and change slowly over time. Furthermore, there are clear transitions from one topic ``state'' to the next, possibly indexing scene transitions in the stimulus. To get a better handle on this temporal structure, we computed a timepoints (1976) by timepoints (1976) correlation matrix of the video model (Fig.~\ref{fig:model}C).  This correlation matrix reveals that the model has a strong, block-diagonal structure with very low off-diagonal correlation values, suggesting the representations of each block are unique and highly descriminable. We found that the ``narrative details'' (a few sentences summarizing the dialogue and specific happenings of the scene) of the video were the most important feature driving this model structure (see Supp. Fig.~\ref{fig:feature-importance}A).

@@

\begin{suppfigure}[t]
\centering
\includegraphics[width=1\textwidth]{figs/supp4_feature_value.pdf}
\caption{\small \textbf{Impact of individual features on topic modeling analysis.} \textbf{A.} Contribution of each feature to model structure. Bars represent the correlation of a video model trained in the absence of a given feature to the model trained on all features. \textbf{B.} Contribution of each feature to video model/recall model relationship. The leftmost bar represents the across-participants mean correlation between the video model and recall models trained on all features. Subsequent bars represent the same relationship between video and recall models trained in the absence of a given feature. Error bars are the standard error of the mean across participants.}
\label{fig:feature-importance}
\end{suppfigure}

>>>>

\subsection*{Modeling the naturalistic stimulus and verbal recall}
We fit a topic model \citep{BleiEtal03} to hand-annotated text descriptions of scenes from the video. The text descriptions contained details of the scene such as the characters present, location, and a short summary of the scene (see Fig.~\ref{fig:schematic} for example text). As depicted in Fig.~\ref{fig:model}A, the topic vectors are sparse and change slowly over time. Furthermore, there are clear transitions from one topic ``state'' to the next, possibly indexing scene transitions in the stimulus. To get a better handle on this temporal structure, we computed a timepoints (1976) by timepoints (1976) correlation matrix of the video model (Fig.~\ref{fig:model}C).  This correlation matrix reveals that the model has a strong, block-diagonal structure with very low off-diagonal correlation values, suggesting the representations of each block are unique and highly discriminable. We found that the ``narrative details'' (a few sentences summarizing the dialogue and specific happenings of the scene) of the video were the most important feature driving this model structure (see Supp. Fig.~\ref{fig:feature-importance}A).

@@

\begin{suppfigure}[t]
\centering
\includegraphics[width=1\textwidth]{figs/supp4_feature_value.pdf}
\caption{\small \textbf{Impact of individual features on topic modeling analysis.} \textbf{A.} Contribution of each feature to model structure. Bars represent the correlation of a video model trained in the absence of a given feature to the model trained on all features. \textbf{B.} Contribution of each feature to video model/recall model relationship. The leftmost bar represents the across-participants mean correlation between the video model and recall models trained on all features. Subsequent bars represent the same relationship between video and recall models trained in the absence of a given feature. Error bars are the standard error of the mean across participants. \textbf{C.}  Individual feature trajectory similarity matrix. The first row/column is the topic trajectory the full video model. Each subsequent row/column is that of a single feature. Shading corresponds to the value of the correlation coefficient (Pearson's $r$) .}
\label{fig:feature-importance}
\end{suppfigure}
